---
title: 'Social distancing data: trajectory and longitudinal clustering'
author: "Aimee Schwab-McCoy"
date: "Last updated: `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
set.seed(041518)
```

## Trajectory clustering using the `traj` package

Data should consist of two data frames:

1. Individual trajectories along the rows
2. Time each individual trajectory was measured (again along the rows)

Trajectory clustering works by computing 24 measures of each trajectory, such as range, mean, standard deviation, slope, etc., and then clustering based on these measures.

```{r, echo=TRUE}
library(traj)

# Individual trajectories are stored in each row
head(example.data$data)

# Time points for each trajectory are stored in the rows
head(example.data$time)

# Compute measures using step1measures
s1 = step1measures(example.data$data, example.data$time, ID = TRUE)

# Use factor analysis to select a subset of features
s2 = step2factors(s1)

# Cluster trajectories based on the factor analysis
s3 = step3clusters(s2, nclusters = 4)

#plot(s3)
#plotMeanTraj(s3)

```

Convert sample code from vignette to data for each state with a protest.

Daily distance difference:

```{r, warning=FALSE, message=FALSE}
County_Data <- read.csv("E:\\ProjectData\\unacast\\sds-v3-full-county.csv")
LockdownProtests <- read.csv("E:\\ProjectData\\unacast\\LockdownProtests.csv")

library(tidyverse)
#library(patchwork)

FIPS_list <- unique(LockdownProtests$FIPS)
State_list <- unique(LockdownProtests$STATE)

County_Data$date <- as.Date(County_Data$date)
LockdownProtests$DATE <- as.Date(LockdownProtests$DATE)

# County_Data <- County_Data %>% 
#   group_by(county_fips) %>%
#   mutate_each_(funs(scale),vars=c('daily_distance_diff',
#                                   'daily_visitation_diff',
#                                   'encounters_rate'))

for(i in 3:3){

    state_data <- County_Data %>% 
      filter(state_code == State_list[i]) %>% 
      select(county_fips, 
             date, 
             daily_distance_diff)
    
    #head(state_data)
    
    state_traj <- state_data %>% 
      pivot_wider(names_from = date, 
                  values_from = daily_distance_diff)
    
    state_date <- state_data %>% 
      select(-daily_distance_diff) %>% 
      pivot_wider(names_from = date, 
                  values_from = date)
    
    s1 = step1measures(state_traj, state_date, ID = TRUE)
    head(s1)
    
    # if(i == 1){ 
    #   discard <- c("m6", "m20", "m21", "m22", "m23", "m24") 
    # } else  if (i==5){ 
    #     discard <- c("m6", "m11", "m12", "m13", "m14", "m15", 
    #                  "m16", "m17", "m19", "m20", "m21", 
    #                  "m22", "m23", "m24")
    # } else {
    #       discard <- "m6"
    #       }
  
    # Use factor analysis to select a subset of features
    s2 = step2factors(s1, discard=c("m3", "m6"))
    
    # Cluster trajectories based on the factor analysis
    s3 = step3clusters(s2, nclusters = 5)
    
    #plot(s3)
    #plotMeanTraj(s3)
    
    clusters <- tibble(county_fips=s3$clusters$ID$county_fips,
                       cluster=s3$clusters$cluster)
    
    state_data <- state_data %>% inner_join(clusters)
    
    plot <- state_data %>% 
        mutate(protest = ifelse(county_fips %in% FIPS_list, 2, 1)) %>%
        arrange(desc(protest)) %>%
        ggplot(aes(x=date, y=daily_distance_diff,
                   group=county_fips)) + 
        geom_line(aes(col=protest, alpha=protest/2)) + 
        facet_grid(~cluster) + 
        guides(col=FALSE, lwd=FALSE, alpha=FALSE) + 
        labs(title=paste('State: ', State_list[i])) 
    
    table <- clusters %>% group_by(cluster) %>% summarize(n=n())
}

plot
table
  
```
