---
title: "Spatio-Temporal Models"
author: "Joshua Eason"
date: "10/16/2020"
output: html_document
---

```{r setup, include=FALSE}

library(tidyverse)
library(stR)
library(plotly)
library(usmap)
library(fitdistrplus)

```

Some additional data wrangling. Making one final time-series dataset which includes only the covid-related features, all time-series observations with the associated coordinates and class designations from the PCA. 
```{r}

no_covid <- read.csv("/mnt/18246CDB246CBCFE/ProjectData/unacast/post-pca.csv")
covid <- read.csv("/mnt/18246CDB246CBCFE/ProjectData/unacast/covid_ts.csv")
dims <- no_covid %>% dplyr::select(FIPS, Dim.1,Dim.2,class2,polyname)
ts_data <- inner_join(covid,dims,by = 'FIPS') %>% dplyr::select(-X)
ts_data$date <- as.Date(ts_data$date)

maps::county.fips %>%
  as_tibble %>% 
  extract(polyname, c("region", "subregion"), "^([^,]+),([^,]+)$") -> dfips
map_data("county") %>% 
  left_join(dfips) -> data
```


Cool looking, but not very helpful...
```{r}

plot_ly(dims, x = ~Dim.1, y = ~Dim.2, z = ~no_covid$Median.Household.Income, 
        marker = list(color = ~no_covid$Median.Household.Income, colorscale = c("Viridis"), showscale = TRUE)) %>%
  add_markers()  %>%
  layout(
    title = 'Counties by PCA Coordinates',
    scene = list(xaxis = list(title = 'Dim 1'),
                 yaxis = list(title = 'Dim 2'),
                 zaxis = list(title = 'MHHI')))
```

## Pre-Covid Models 
Fitting a linear regression model to see which variables are significant for predicting the metrics we care about before covid

Creating a subset based on a single day (2/24/2020)
```{r}
day <- ts_data %>% filter(date == as.Date("2020-02-24"))
features <- day %>% dplyr::select(-c("FIPS","covid","date","county_name","state_code","polyname","week"))
vis_only <- features %>% dplyr::select(-c("daily_distance_diff","encounters_rate"))
dist_only <- features %>% dplyr::select(-c("daily_visitation_diff","encounters_rate"))
encounter_only <- features %>% dplyr::select(-c("daily_distance_diff","daily_visitation_diff"))
```

Basic LM - each of the three for 2/24/2020
```{r}
#Visitation only
model_vis <- lm(daily_visitation_diff ~., data = vis_only)
summary(model_vis)

#Distance only
model_dist <- lm(daily_distance_diff ~., data = dist_only)
summary(model_dist)

#Encounter rate only
model_encounter <- lm(encounters_rate ~., data = encounter_only)
summary(model_encounter)
```

Creating a subset based on a single day (3/08/2020)
```{r}
day <- ts_data %>% filter(date == as.Date("2020-03-08"))
features <- day %>% dplyr::select(-c("FIPS","covid","date","county_name","state_code","polyname","week"))
vis_only <- features %>% dplyr::select(-c("daily_distance_diff","encounters_rate"))
dist_only <- features %>% dplyr::select(-c("daily_visitation_diff","encounters_rate"))
encounter_only <- features %>% dplyr::select(-c("daily_distance_diff","daily_visitation_diff"))
```

Basic LM - each of the three for 3/08/2020
```{r}
#Visitation only
model_vis <- lm(daily_visitation_diff ~., data = vis_only)
summary(model_vis)

#Distance only
model_dist <- lm(daily_distance_diff ~., data = dist_only)
summary(model_dist)

#Encounter rate only
model_encounter <- lm(encounters_rate ~., data = encounter_only)
summary(model_encounter)
```


## Post-Covid Models
Here I fitted the same linear regression on each of the three adherence metrics to see which variables are significant for two randomly selected dates post covid...

Creating a subset based on a single day (4/1/2020)
```{r}
day <- ts_data %>% filter(date == as.Date("2020-04-01"))
features <- day %>% dplyr::select(-c("FIPS","covid","date","county_name","state_code","polyname","week"))
vis_only <- features %>% dplyr::select(-c("daily_distance_diff","encounters_rate"))
dist_only <- features %>% dplyr::select(-c("daily_visitation_diff","encounters_rate"))
encounter_only <- features %>% dplyr::select(-c("daily_distance_diff","daily_visitation_diff"))
```

Basic LM - each of the three for 4/01/2020
```{r}
#Visitation only
model_vis <- lm(daily_visitation_diff ~., data = vis_only)
summary(model_vis)

#Distance only
model_dist <- lm(daily_distance_diff ~., data = dist_only)
summary(model_dist)

#Encounter rate only
model_encounter <- lm(encounters_rate ~., data = encounter_only)
summary(model_encounter)
```

Creating a subset based on a single day (5/17/2020)
```{r}
day <- ts_data %>% filter(date == as.Date("2020-05-17"))
features <- day %>% dplyr::select(-c("FIPS","covid","date","county_name","state_code","polyname","week"))
vis_only <- features %>% dplyr::select(-c("daily_distance_diff","encounters_rate"))
dist_only <- features %>% dplyr::select(-c("daily_visitation_diff","encounters_rate"))
encounter_only <- features %>% dplyr::select(-c("daily_distance_diff","daily_visitation_diff"))
```

Basic LM - each of the three for 5/17/2020
```{r}
#Visitation only
model_vis <- lm(daily_visitation_diff ~., data = vis_only)
summary(model_vis)

#Distance only
model_dist <- lm(daily_distance_diff ~., data = dist_only)
summary(model_dist)

#Encounter rate only
model_encounter <- lm(encounters_rate ~., data = encounter_only)
summary(model_encounter)
```

## What are the distributions of the response variables for each of the days?

Here, it looks like I may have 3 different distributions of the response variables, this suggests that I may need to consider different formula families for the GLM to model them properly. It may also be the case that they are all poisson distributed. 
```{r}
#visualizing the distribution of the response variables to help decide the distribution

#Normal-ish? Higher lambda
vis_only_nontrivial <- vis_only %>% filter(daily_visitation_diff != 0) 
vis_only_nontrivial %>% ggplot(aes(x = daily_visitation_diff)) + geom_histogram(bins = 50)

#Skew right - Lower lambda
dist_only_nontrivial <- dist_only %>% filter(daily_distance_diff != 0) 
dist_only_nontrivial %>% ggplot(aes(x = daily_distance_diff)) + geom_histogram(bins = 50)

#Very skew right - very low lambda
encounter_only_nontrivial <- encounter_only %>% filter(encounters_rate != 0) 
encounter_only_nontrivial %>% ggplot(aes(x = log(encounters_rate))) + geom_histogram(bins = 50)

```


```{r}
#dailies are both "normal"
#encounters to log-normal
dist <- descdist(vis_only$daily_visitation_diff,  discrete = FALSE)
ft <- fitdist(vis_only$daily_visitation_diff, distr = "norm" )
ft
denscomp(ft)

```


This still doesnt work...
```{r}
#need to remove the dimensions from the glm to run the simple version.

#https://datascienceplus.com/spatial-regression-in-r-part-1-spamm-vs-glmmtmb/

vis_only <- vis_only %>% dplyr::select(-c("Dim.1","Dim.2","class2"))

vis_glm <- glm(vis_only$daily_visitation_diff~., family=gaussian("identity"), data= vis_only)
summary(vis_glm)

```

# Grouping by week
```{r}
library(dplyr)
ts_week <- ts_data %>% dplyr::group_by(FIPS,week) %>% dplyr::summarize(dist = mean(daily_distance_diff),
                                                    vis = mean(daily_visitation_diff),
                                                    encounter = mean(encounters_rate),
                                                    cases = max(cases),
                                                    deaths = max(deaths),
                                                    week = max(week),
                                                    death_rate = max(death_rate),
                                                    death_per_capita = max(death_per_capita),
                                                    dim1 = max(Dim.1),
                                                    dim2= max(Dim.2))

```

# Fitting a model to the time series data

```{r}
library(lme4)
model <- lmer(encounter ~ week + cases + deaths + death_rate + death_per_capita + (1|dim1) + (1|dim2),data=ts_week)
summary(model)
```


